---
title: "BBets"
author: "Jovygomez"
output: html_document
date: "2024-09-27"
---


This assignment has three questions:  

1. in a class of 35 students, Dr. J bets $1 against student's $1  that at least 2 students have the same birthday, month, day.
Find the probability thatat least two student have the same birthday.  If 26 students in the class take the bet how much is Dr. J expected to win/lose.   How much should Dr. J wager against the student's $1 to make this a fair bet?

2. Assume class size is 26 instead of 35.  With the same bet ($1 ot $1) what is the probability Dr. J wins the bet now.  If 20 of the 26 students make the bet, what is Dr. J's expected profit/loss?

3. What size class (how many students) makes the $1 againt $1 a fair bet?

Question 1: Answer  Below 


```{r, echo=TRUE}
# Set seed for reproducibility
set.seed(123)


# Function to simulate one birthday experiment
birthday_simulation <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  return(any(duplicated(birthdays)))
}

# Initialize parameters
class_size <- 35
num_simulations <- 100
probabilities <- numeric(num_simulations)

# Run simulations
for (i in 1:num_simulations) {
  success_count <- sum(replicate(1000, birthday_simulation(class_size)))
  probabilities[i] <- success_count / 1000
}

# Plot the histogram of probabilities
hist(probabilities, main = "Histogram of Probabilities (Birthday Problem)", 
     xlab = "Probability", col = "lightblue", border = "black")

# Calculate mean of the probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")

# Calculate 95% confidence interval
conf_interval <- quantile(probabilities, c(0.025, 0.975))
cat("95% Confidence Interval =", conf_interval, "\n")
```

```{r}
EV= mean_prob*26-(1-mean_prob)*26
print(paste("ev=",EV))
```

```{r}
FAIR=(mean_prob)/(1-mean_prob)
print(paste("fairbet=",FAIR))
```

Find the probability that at least two student have the same birthday.  If 26 students in the class take the bet how much is Dr. J expected to win/lose. 
-He is expected to win

How much should Dr. J wager against the student's $1 to make this a fair bet?

-Dr. J would have to bet 4.6 dollars for it to be a fair bet. 

##Question 2
```{r, echo=TRUE}
# Set seed for reproducibility
set.seed(123)


# Function to simulate one birthday experiment
birthday_simulation <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  return(any(duplicated(birthdays)))
}

# Initialize parameters
class_size <- 26
num_simulations <- 100
probabilities <- numeric(num_simulations)

# Run simulations
for (i in 1:num_simulations) {
  success_count <- sum(replicate(1000, birthday_simulation(class_size)))
  probabilities[i] <- success_count / 1000
}

# Plot the histogram of probabilities
hist(probabilities, main = "Histogram of Probabilities (Birthday Problem)", 
     xlab = "Probability", col = "lightblue", border = "black")

# Calculate mean of the probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")

# Calculate 95% confidence interval
conf_interval <- quantile(probabilities, c(0.025, 0.975))
cat("95% Confidence Interval =", conf_interval, "\n")
```

```{r}
EV= mean_prob*20-(1-mean_prob)*20
print(paste("ev=",EV))
```

```{r}
FAIR=(mean_prob)/(1-mean_prob)
print(paste("fairbet=",FAIR))
```

Assume class size is 26 instead of 35. With the same bet ($1 to $1) what is the probability that Dr. J wins now? 
If 20 of the 26 students make the bet, what is Dr. J expected profit/loss? 

##Question 3

```{r, echo=TRUE}
# Set seed for reproducibility
set.seed(123)


# Function to simulate one birthday experiment
birthday_simulation <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  return(any(duplicated(birthdays)))
}

# Initialize parameters
class_size <- 23
num_simulations <- 100
probabilities <- numeric(num_simulations)

# Run simulations
for (i in 1:num_simulations) {
  success_count <- sum(replicate(1000, birthday_simulation(class_size)))
  probabilities[i] <- success_count / 1000
}

# Plot the histogram of probabilities
hist(probabilities, main = "Histogram of Probabilities (Birthday Problem)", 
     xlab = "Probability", col = "lightblue", border = "black")

# Calculate mean of the probabilities
mean_prob <- mean(probabilities)
cat("Mean =", mean_prob, "\n")

# Calculate 95% confidence interval
conf_interval <- quantile(probabilities, c(0.025, 0.975))
cat("95% Confidence Interval =", conf_interval, "\n")
```

```{r}
EV= mean_prob*20-(1-mean_prob)*20
print(paste("ev=",EV))
```

In order for it to be a Fair bet the class would need to have 23 students.
